---
title: "SUSA - A PrimeR"
author: "Nicholas Lai"
date: "September 6, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, cache = TRUE)

#install.packages(c('ggplot2', 'dplyr', 'viridis'), repos = 'https://cran.cnr.berkeley.edu/')

library(ggplot2) #data visualization framework
library(dplyr) #tidy data
library(viridis) #Colorblind-friendly gradients
```

One of the most common uses of R is to generate beautiful data visualizations. In fact, many professional data journalism outfits such as FiveThirtyEight and The New York Times use R to create their graphics. 

Today, we will be creating a graph of the United States by county. This may sound like a daunting first task, but because of the wonderful packages that the R community has produced, it will become clear that doing so is quite managable. 

#Getting the Map

The ggplot2 R package is one of the most powerful graphical tools available in R. One of the functions of the package allows us to extract map data for the US and for counties of the US.

```{r map}
usa <- map_data("usa")
counties <- map_data("county")
```

We can read in census data using functions in R, and using the head function, we can inspect the data's structure. 

```{r data}
county_data <- read.csv("acs2015_county_data.csv")
head(county_data)
```

In order to plot the data on a map, we need to merge the data with the map data. 

```{r cleaning}
county_data$County <- sapply(county_data$County, tolower)
head(counties)
colnames(county_data)[colnames(county_data)=="County"] <- "subregion"

usa_county_data <- inner_join(counties, county_data, by = "subregion")

head(usa_county_data)
```

After we have the merged data, we are ready to plot!

```{r plotting}
ggplot(data = usa, aes(x=long, y = lat, group = group)) + geom_polygon(fill = 'gray')+
  coord_fixed(1.3) + geom_polygon(data = usa_county_data, aes(fill = White), color = "white") + scale_fill_viridis(option="magma")
```

#Your Turn

Say you are interested in the percentage of people in a given county that are citizens. To get this data, divide the column of 'Citizen' with the column of 'TotalPop'.

```{r}
usa_county_data$prop_citizen <- usa_county_data$Citizen/usa_county_data$TotalPop

head(usa_county_data) 
```

Say you are interested in total number of white people in a county. How would you transform your data to get those values?

```{r}
#Your code here
```

Now, plot one of the two above new variables to the US county map like above.

```{r}
#Your code here
```

Look back at the usa_county_data object. It has a very redundant column as a quirk of our 'inner.join'. Which columns are redundant?

We can remove it:
```{r}
#usa_county_data$col_name <- NULL
```

Challenge Question:

In a similar way to which we can visualize the county_data on the county level, we can visualize it on a state level. 

Aggregate a variable of the data by state and join that to a map_data object, and plot that variable on the state level using a modification of the above method.

```{r}
#Your code here
```

