---
title: "SUSA - A PrimeR"
author: "Nicholas Lai"
date: "September 6, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, cache = TRUE)

#install.packages(c('ggplot2', 'dplyr', 'viridis'), repos = 'https://cran.cnr.berkeley.edu/')

library(ggplot2) #data visualization framework
library(dplyr) #tidy data
library(viridis) #Colorblind-friendly gradients
```
__Welcome to R!__

Let's begin with some R basics.

#Vectorization

When it comes to vector arithmetic, R is pretty straightforward because it uses element-wise arithmetic. One special aspect of R is that it is conducive to something called _vector recycling_. 

For example, say we have some temperatures in Celsius

```{r}
alpha <- c(6, 420, 89, 3429990, 666, -294)
```
and we want to convert to temperatures in Kelvin. We know $Kelvin = Celsius + 273.15$. 

R makes this easy to do:

```{r}
alpha + 273.15 #adds 273.15 to each component in alpha
```
Vector recycling is iterative. Look at what the following statements return:

```{r}
c(6, 5, 4, 3, 2, 1) + c(-3,2)
```
The command above is interpreted as adding values of a secondary vector to each subsequent element in the first vector. Since the vectors are different sizes we just add the values of the smaller length vector repeatedly until we have added a value to each element of the larger length vector.

```{r}
c(5, 4, 3, 2, 1) + c(-3,2)
```
The command above tries the same logic but it cannot fully parse through the smaller length vector an integer number of times. Thus we get the warning prompt.

#Your turn

What is an easy way to manipulate gamma into delta (what vector arithmetic can be performed on alpha to get beta)?
```{r}
gamma <- c(45, 2, 98, 33, 67, 1)
delta <- c(-3, 10, 40, 41, 19, 9)
#Your code here
```

__Vector Coercion__

Vectors are atomic structures (meaning their components must all be of the same type). If you mix data types, vectors undergo coercion. So be careful to only enter values of the same datatype when making a vector.

Suppose we mix logical components with numeric ones:

```{r}
v <- c(TRUE, FALSE, 56, 72)
v
```
Then the vector becomes numeric because it must choose a single datatype.

Suppose we mix characters with numeric:

```{r}
w <- c("bone", "apple", "tea", 4, 2, 0)
w
```

Then the vector becomes a string vector. What happens if we mix logicals (TRUE, FALSE) with characters ("strings")?

```{r}
#You can make a guess or try out some code here to see for yourself.
```

#Formatting Equations with LaTex Syntax

Let's derive the quadratic formula: $$x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$$

We start with a quadratic equation and look for its roots: $$ax^2 + bx + c = 0$$

Now multiply both sides by $4a$: 
$Your code here$

Then add $b^2 - 4ac$ to both sides:
$Your code here$

Then factor the left hand side into a square:
$$(2ax + b)^2 = b^2 - 4ac$$

Then take the square root of each side:
Hint: $\pm$ gives a plus or minus sign, $\sqrt{w}$ gives square root
$Your code here$

Then subtract $b$ from each side:
$$2ax = -b \pm \sqrt{b^2 - 4ac}$$

Then solve for x:
Hint: $\frac{u}{v}$ gives a pretty fraction.
$Your code here$ And you're done!

One of the most common uses of R is to generate beautiful data visualizations. In fact, many professional data journalism outfits such as FiveThirtyEight and The New York Times use R to create their graphics. 

Today, we will be creating a graph of the United States by county. This may sound like a daunting first task, but because of the wonderful packages that the R community has produced, it will become clear that doing so is quite managable. 

#Getting the Map

The ggplot2 R package is one of the most powerful graphical tools available in R. One of the functions of the package allows us to extract map data for the US and for counties of the US.

```{r map}
usa <- map_data("usa")
counties <- map_data("county")
```

We can read in census data using functions in R, and using the head function, we can inspect the data's structure. 

```{r data}
county_data <- read.csv("acs2015_county_data.csv")
head(county_data)
```

In order to plot the data on a map, we need to merge the data with the map data. 

```{r cleaning}
county_data$County <- sapply(county_data$County, tolower)
head(counties)
colnames(county_data)[colnames(county_data)=="County"] <- "subregion"

usa_county_data <- inner_join(counties, county_data, by = "subregion")

head(usa_county_data)
```

After we have the merged data, we are ready to plot!

```{r plotting}
ggplot(data = usa, aes(x=long, y = lat, group = group)) + geom_polygon(fill = 'gray')+
  coord_fixed(1.3) + geom_polygon(data = usa_county_data, aes(fill = White), color = "white") + scale_fill_viridis(option="magma")
```

#Your Turn

Say you are interested in the percentage of people in a given county that are citizens. To get this data, divide the column of 'Citizen' with the column of 'TotalPop'.

```{r}
usa_county_data$prop_citizen <- usa_county_data$Citizen/usa_county_data$TotalPop

head(usa_county_data) 
```

Say you are interested in total number of white people in a county. How would you transform your data to get those values?

```{r}
#Your code here
```

Now, plot one of the two above new variables to the US county map like above.

```{r}
#Your code here
```

Look back at the usa_county_data object. It has a very redundant column as a quirk of our 'inner.join'. Which columns are redundant?

We can remove it:
```{r}
#usa_county_data$col_name <- NULL
```

Challenge Question:

In a similar way to which we can visualize the county_data on the county level, we can visualize it on a state level. 

Aggregate a variable of the data by state and join that to a map_data object, and plot that variable on the state level using a modification of the above method.

```{r}
#Your code here
```


